---
title: "Raster HW"
author: Kt Nowak
format: 
  html:
    code-tools: true
    toc: false
    echo: true
---

Rasters are spatial continuous distributions of geospatial data.  We have been shown that we an work with data in raster format using either `raster` and working on object directly, or via `dplyr` and using the normal data manipulation routines that we've learned for non-spatial data.  In this *in class* exercise, you may use both of these techniques to load, manipulate, and display geospatial data.

## The Raw Data

The data for this will be the same Baja California data we showed in the lectures on [points](https://dyerlab.github.io/ENVS-Lectures/spatial/spatial_points/slides.html#1) and [rasters](https://dyerlab.github.io/ENVS-Lectures/spatial/rasters/slides.html#1).  

```{r}
raster_url <- "https://github.com/DyerlabTeaching/Raster-Data/raw/main/data/alt_22.tif"
beetle_url <- "https://raw.githubusercontent.com/DyerlabTeaching/Raster-Data/main/data/AraptusDispersalBias.csv"
```


## The Questions

1. Load the raster and point data in and crop to an appropriate size to display the locations of the sampling plots and make a label for each site.

```{r}
library( tidyverse )
library( raster )
library( sf ) 
library(dplyr)
library(ggspatial)

elev <- raster( raster_url )

read_csv( beetle_url )  %>% 
  st_as_sf( coords = c("Longitude","Latitude"), 
            crs = 4326 ) -> beetles 

beetles
```



```{r}
plot( elev, xlab="Longitude" )
```


```{r}
beetles %>%
  st_union() %>%
  st_buffer( dist = 1 ) %>%
  st_bbox() 

baja_extent <- extent( c(-115.29354,-108,22,30))

baja_elev <- crop( elev, baja_extent )
baja_elev
plot( baja_elev) 

```


```{r}
plot( baja_elev, xlab="Longitude", ylab="Latitude")
#text( beetles, add=TRUE, pch=16, col="red" )
```


```{r}
library( ggrepel )
baja_elev %>%
  rasterToPoints() %>%
  as.data.frame() %>% 
  ggplot() +
  geom_raster( aes(x,y,fill=alt_22) ) +
  geom_sf_label( aes(label=Site), 
                 data = beetles) +
  coord_equal()
```




2. Use the click() function to crop the raster and filter the sites to include only the sampling locations that are in the Cape regions (e.g., south of site labeled #75).  Plot the raster of elevation and all cape region locales with sex-ratio represented as the size of the plot symbol.

```{r}

site_75 <- beetles[beetles$Site == 75, ]

southernmost_lat <- st_coordinates(site_75)[1, 2]

extent_south <- extent(xmin(elev), xmax(elev), -Inf, southernmost_lat)

elevation_cropped <- crop(elev, extent_south)

extent_poly <- as(extent(elevation_cropped), "SpatialPolygons")

extent_poly_sf <- st_as_sf(extent_poly)

beetles_sf <- st_as_sf(beetles, coords = c("longitude", "latitude"), crs = 4326)

st_crs(beetles_sf) <- st_crs(extent_poly_sf)

cape_sites <- st_intersection(beetles_sf, extent_poly_sf)

elevation_values <- values(elevation_cropped)

cape_map <- elev + 
  scale_fill_gradient2(low = "darkolivegreen", mid = "yellow", high = "brown", midpoint = 1000) +
  geom_sf(data = cape_sites, aes(size = MFRatio), alpha = 0.75) + 
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", style = north_arrow_fancy_orienteering()) +
  ggtitle("Site Map for Cape Region") +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal() +
  geom_text_repel(aes(label = Site,
                      geometry = geometry),
                  data = cape_sites,
                  stat = "sf_coordinates", 
                  size = 4, 
                  color = "black") + 
  coord_sf()


ggplot() +
  geom_raster(data = ( xy = TRUE), 
               aes(x = x, y = y, fill = values)) +
  geom_point(data = beetles, aes(x = longitude, y = latitude, size = MFRatio, color = SexRatio)) +
  scale_fill_gradient(name = "Elevation") +
  scale_size(name = "Sex Ratio") +
  labs(title = "Elevation and Site Locations", x = "Longitude", y = "Latitude")



print(cape_map)

```

3. The peninsula of Baja California is divided into the States of Baja California Norte and Baja California Sur. The border between these states is at 28° Latitude. Divide the sample locations into groups based on which state they are located in and plot the average sex ratio of the sample sites partitioned by each site.  

```{r}

beetles <- beetles %>%
  mutate(
    longitude = str_extract(geometry, "-?\\d+\\.\\d+"),
    latitude = str_extract(geometry, "(?<= )-?\\d+\\.\\d+")
  ) %>%
  select(-geometry)

new_beetles <- beetles %>%
  mutate(State = ifelse(latitude > 28, "Baja California Norte", "Baja California Sur"))


```

4. Is there a relationship between the observed sex ratio and the elevation at that site?  Plot these, and in the text, perform a correlation test (cor.test()) and report the values as if you were writing them up in a results section.

