---
title: "Raster HW"
author: Kt Nowak
format: 
  html:
    code-tools: true
    toc: false
---

Rasters are spatial continuous distributions of geospatial data.  We have been shown that we an work with data in raster format using either `raster` and working on object directly, or via `dplyr` and using the normal data manipulation routines that we've learned for non-spatial data.  In this *in class* exercise, you may use both of these techniques to load, manipulate, and display geospatial data.

## The Raw Data

The data for this will be the same Baja California data we showed in the lectures on [points](https://dyerlab.github.io/ENVS-Lectures/spatial/spatial_points/slides.html#1) and [rasters](https://dyerlab.github.io/ENVS-Lectures/spatial/rasters/slides.html#1).  

```{r}
raster_url <- "https://github.com/DyerlabTeaching/Raster-Data/raw/main/data/alt_22.tif"
beetle_url <- "https://raw.githubusercontent.com/DyerlabTeaching/Raster-Data/main/data/AraptusDispersalBias.csv"
```


## The Questions

1. Load the raster and point data in and crop to an appropriate size to display the locations of the sampling plots and make a label for each site.

```{r}
library( tidyverse )
library( raster )
library( sf ) 

elev <- raster( raster_url )

read_csv( beetle_url )  %>% 
  st_as_sf( coords = c("Longitude","Latitude"), 
            crs = 4326 ) -> beetles 

beetles
```



```{r}
plot( elev, xlab="Longitude" )
```


```{r}
beetles %>%
  st_union() %>%
  st_buffer( dist = 1 ) %>%
  st_bbox() 

baja_extent <- extent( c(-115.29354,-108,22,30))

baja_elev <- crop( elev, baja_extent )
baja_elev
plot( baja_elev) 

```


```{r}
plot( baja_elev, xlab="Longitude", ylab="Latitude")
text( beetles, add=TRUE, pch=16, col="red" )
```


```{r}
library( ggrepel )
baja_elev %>%
  rasterToPoints() %>%
  as.data.frame() %>% 
  ggplot() +
  geom_raster( aes(x,y,fill=alt_22) ) +
  geom_sf_label( aes(label=Site), 
                 data = beetles) +
  coord_equal()
```




2. Use the click() function to crop the raster and filter the sites to include only the sampling locations that are in the Cape regions (e.g., south of site labeled #75).  Plot the raster of elevation and all cape region locales with sex-ratio represented as the size of the plot symbol.

```{r}

site75 -> 

cape_map <- elev + 
  scale_fill_gradient2(low = "darkolivegreen", mid = "yellow", high = "brown", midpoint = 1000) +
  geom_sf(data = cape_sites, aes(size = MFRatio), color = "dodgerblue2", alpha = 0.75) + 
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", style = north_arrow_fancy_orienteering()) +
  ggtitle("Site Map for Cape Region") +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal() +
  geom_text_repel(aes(label = Site,
                      geometry = geometry),
                  data = cape_sites,
                  stat = "sf_coordinates", 
                  size = 4, 
                  color = "black") + 
  coord_sf()



print(cape_map)

```







3. The peninsula of Baja California is divided into the States of Baja California Norte and Baja California Sur. The border between these states is at 28° Latitude. Divide the sample locations into groups based on which state they are located in and plot the average sex ratio of the sample sites partitioned by each site.  

4. Is there a relationship between the observed sex ratio and the elevation at that site?  Plot these, and in the text, perform a correlation test (cor.test()) and report the values as if you were writing them up in a results section.

