---
title: "Spatial Data HW"
author: KT Nowak
format: html
code-tools: true
warning: FALSE
---
<center>
![*Araptus attenuatus* the enigmatic Sonoran Desert Bark
Beetle.](https://live.staticflickr.com/65535/50441339417_74e04216fa_w_d.jpg)
</center>


## The Data - Preprocessing
The data for this is some site-level data from the Sonoran Desert bark beetle. The
URL is in the next chunk, and is currently being hosted on my class [Github
repository](https://github.com/dyerlab/ENVS-Lectures). Load it in.
```{r}
url <- "https://raw.githubusercontent.com/dyerlab/ENVS-Lectures/master/data/Araptus_Disperal_Bias.csv"

beetle <- read.csv(url)

```
# Bringing in Packages

```{r}
library(tidyverse)

library(dplyr)

library(knitr)

library(kableExtra)

library(lubridate)

library(leaflet)

library(leaflet.providers)

library(sf)
```
### Interactive Map

Use the `leaflet` library and make an interactive map. If you can, add a label to
each marker with the ID of the site. You can look at the provider tiles [here]
(https://leaflet-extras.github.io/leaflet-providers/preview/) and use on in the
`addProviderTiles()` function.

There is a great tutorial [here](https://rstudio.github.io/leaflet/map_widget.html)
on how to customize the leaflet display. You may want to go look at it and see
what you can use to make a better display.

```{r}
# make a quick leaflet map here to get an idea of

beetle %>%
  mutate( Label = paste( "Site:", Site, 
                         "<hr>\nFemales:", Females, 
                         "<br>Males: ", Males,
                         "<br>Suitability:", Suitability) ) -> df

library( leaflet )
df %>%
  leaflet() %>%
  addMarkers( ~Longitude, ~Latitude, popup = ~Label ) %>%
  addProviderTiles( "OpenTopoMap" )

```
### Converting to `sf` Objects

Convert the spatial data in the beetle data.frame into an `sf` object. Make sure
you set the Coordinate Reference System to $4326$.
```{r}
# convert raw data to simple features in your data.frame

beetle %>% 
  st_as_sf( coords=c("Longitude","Latitude"), crs=4326 ) -> beetle


head(beetle)
```
### Questions
For these questions, you may want to look at the `sf` cheat sheet
[here](https://github.com/rstudio/cheatsheets/raw/main/sf.pdf) to find the
appropriate geospatial function.

1. How far apart (in km) are the sites Constitución (`const`) and San Francisquito
(`sfran`)?

```{r}

```

2. Is there a correlation between habitat suitability and the ratio of males to
female beetles sampled at each locale?
```{r}

```

3. Make a plot using `ggplot` of the data with suitability as the size of the
points and the sex ratio as the fill color.
```{r}

```
4. Reproject the data and plot it again using an EPSG code defined for Mexico UTM
zone 12N (see [epgs.io](https://epsg.io) for projections).
```{r}
#| echo: false
cor.response <- cor.test( iris$Petal.Length, iris$Sepal.Width)
```
We failed to find a significant relationship between sepal length and width
(Pearson Correlation, $\rho$ = `r cor.response$estimate`, P = `r
cor.response$p.value`).