---
title: "Spatial Data HW"
author: KT Nowak
format: html
code-tools: true
warning: FALSE
---
<center>
![*Araptus attenuatus* the enigmatic Sonoran Desert Bark
Beetle.](https://live.staticflickr.com/65535/50441339417_74e04216fa_w_d.jpg)
</center>


## The Data - Preprocessing
The data for this is some site-level data from the Sonoran Desert bark beetle. The
URL is in the next chunk, and is currently being hosted on my class [Github
repository](https://github.com/dyerlab/ENVS-Lectures). Load it in.
```{r}
url <- "https://raw.githubusercontent.com/dyerlab/ENVS-Lectures/master/data/Araptus_Disperal_Bias.csv"

beetle <- read.csv(url)

```
# Bringing in Packages

```{r}
library(tidyverse)

library(dplyr)

library(knitr)

library(kableExtra)

library(lubridate)

library(leaflet)

library(leaflet.providers)

library(sf)
```
### Interactive Map

Use the `leaflet` library and make an interactive map. If you can, add a label to
each marker with the ID of the site. You can look at the provider tiles [here]
(https://leaflet-extras.github.io/leaflet-providers/preview/) and use on in the
`addProviderTiles()` function.

There is a great tutorial [here](https://rstudio.github.io/leaflet/map_widget.html)
on how to customize the leaflet display. You may want to go look at it and see
what you can use to make a better display.

```{r}
# make a quick leaflet map here to get an idea of

beetle %>%
  mutate( Label = paste( "Site:", Site, 
                         "<hr>\nFemales:", Females, 
                         "<br>Males: ", Males,
                         "<br>Suitability:", Suitability) ) -> df

library( leaflet )
df %>%
  leaflet() %>%
  addMarkers( ~Longitude, ~Latitude, popup = ~Label ) %>%
  addProviderTiles( "OpenTopoMap" )

```
### Converting to `sf` Objects

Convert the spatial data in the beetle data.frame into an `sf` object. Make sure
you set the Coordinate Reference System to $4326$.
```{r}
# convert raw data to simple features in your data.frame

beetle %>% 
  st_as_sf( coords=c("Longitude","Latitude"), crs=4326 ) -> beetle


head(beetle)
```
### Questions
For these questions, you may want to look at the `sf` cheat sheet
[here](https://github.com/rstudio/cheatsheets/raw/main/sf.pdf) to find the
appropriate geospatial function.

1. How far apart (in km) are the sites Constitución (`const`) and San Francisquito
(`sfran`)?

The 2 sites are 2.670225 meters apart.

```{r}

const <- st_as_sf(data.frame(latitude = 25.02470, longitude = -111.6750), coords = c("longitude", "latitude"))

sfran <- st_as_sf(data.frame(latitude = 27.36320, longitude = -112.9640), coords = c("longitude", "latitude"))

distance <- st_distance(const, sfran)
print(distance)
```

2. Is there a correlation between habitat suitability and the ratio of males to
female beetles sampled at each locale?

There is a moderately significant negative correlation. 

```{r}

correlation_coefficient <- cor.test(df$Suitability, df$MFRatio)

print(correlation_coefficient)



```

3. Make a plot using `ggplot` of the data with suitability as the size of the
points and the sex ratio as the fill color.

```{r}

ggplot(df, aes(x = df$Latitude, y = df$Longitude,fill = df$MFRatio, size = df$Suitability)) +
  scale_fill_gradient2(name = "Sex Ratio", limits = c(0, 3)) +
  geom_point(shape = 21) + 
  scale_size(name = "Suitability") +
  labs(title = "Suitability and Male to Female Ratio of the Sonoran Desert Bark Beetle ",
       x = "Latitude",
       y = "Longitude")

```


4. Reproject the data and plot it again using an EPSG code defined for Mexico UTM
zone 12N (see [epgs.io](https://epsg.io) for projections).
```{r}

sf_df <- st_as_sf(df, coords = c("Longitude", "Latitude"), crs = 4326) 

sf_df_utm <- st_transform(sf_df, crs = 32613)

sf_df_utm <- df %>%
  mutate(MFRatio_category = cut(MFRatio, breaks = c(-Inf, 0.5, 1, 2, Inf), 
                               labels = c("Low", "Medium", "High", "Very High")))


ggplot(sf_df_utm, aes(x = Longitude, y = Latitude, fill = MFRatio_category, size = Suitability)) +
  geom_point(shape = 21) +
  scale_fill_discrete(name = "Sex Ratio Category") +
  scale_size(name = "Suitability") +
  labs(title = "Suitability and Sex Ratio of the Sonoran Desert Bark Beetle ",
       x = "Longitude",
       y = "Latitude")

#| echo: false
cor.response <- cor.test( iris$Petal.Length, iris$Sepal.Width)
```
We failed to find a significant relationship between sepal length and width
(Pearson Correlation, $\rho$ = `r cor.response$estimate`, P = `r
cor.response$p.value`).